---
title: "Search trends analysis"
author: "Alex Papageorgiou"
date: "15/2/2022"
output: github_document
---



# Setup

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(readr)
library(lubridate)
library(textstem)
library(reticulate)
library(zoo)
library(broom)
library(scales)
library(ggrepel)
library(gghighlight)
library(ggthemes)



source("gg-functions.R")

# theme_set(theme_light())

# font =  "Roboto Condensed"



options(scipen = 999)
options(digits = 6)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(extrafont)
loadfonts(device = "win")
```


# Data

```{r paged.print=TRUE}
search_data <- read_rds("startup_search_data.rds")


search_data %>% sample_n(20)
```



# STATS


## Add months columns

```{r}

months_df <- tibble(month = ymd("2018-01-01") + months(0:47))


sd_long <- search_data %>% unnest_longer(searches_past_months)


sd_monthly <- sd_long %>%
  mutate(
    month = rep(months_df$month, len = nrow(sd_long)),
    month_counter = rep(1:48, len = nrow(sd_long))
  ) %>%
  mutate(year = year(month) + yday(month) / 365) %>%
  select(Keyword = keyword, seed, searches_past_months, avg_monthly_searches, month_counter, month, year)
```


## data to csv

```{r}
# sd_monthly %>%
#   select(keyword = Keyword, seed, month, search_volume = searches_past_months) %>%
#   write_csv("startup_search_data.csv")
```


## 3m Rolling avg

```{r}

sd_month_filt <- sd_monthly %>%
  group_by(Keyword) %>%
  arrange(desc(year)) %>%
  dplyr::filter(max(searches_past_months) > 30) %>%
  ungroup()

rolling <- sd_month_filt %>%
  group_by(Keyword) %>%
  mutate(n = n()) %>%
  arrange(year) %>%
  mutate(roll_avg = zoo::rollmean(searches_past_months, 3, fill = NA)) %>%
  mutate(roll_avg_round = round(roll_avg, digits = 0)) %>%
  ungroup()
```

## Lms

to estimate avg monthly change

```{r}
models <- rolling %>% # mutate(search_volume =ifelse(search_volume <100, NA , search_volume) )
  group_by(Keyword) %>%
  # dplyr::filter(cumany(searches_past_months>0)) %>%
  nest() %>%
  mutate(model = map(data, function(df) lm(roll_avg_round ~ month_counter, data = df))) %>%
  mutate(
    by_term = map(model, broom::tidy),
    perf = map(model, broom::glance)
  )

# anti <- rolling %>% anti_join(models, by=c('Keyword', "searches_past_months"))

by_term <- models %>%
  unnest(by_term) %>%
  select(Keyword, term, estimate)

perf <- models %>%
  unnest(perf) %>%
  select(Keyword, r.squared)


by_term_perf <- by_term %>%
  left_join(perf, by = "Keyword") %>%
  arrange(desc(r.squared)) %>%
  dplyr::filter(term == "month_counter")


kw_rsq_vol <- by_term_perf %>%
  left_join(search_data %>% select(Keyword = keyword, avg_monthly_searches)) %>%
  relocate(avg_monthly_searches, .after = Keyword) %>%
  mutate(estimate = round(estimate, digits = 2))
```

## Natural log model

to estimate growth rate

```{r}
model_ln <- rolling %>% # mutate(search_volume =ifelse(search_volume <100, NA , search_volume) )
  dplyr::filter(roll_avg_round > 0) %>%
  group_by(Keyword) %>%
  nest() %>%
  mutate(model = map(data, function(df) lm(log(roll_avg_round) ~ month_counter, data = df))) %>%
  mutate(
    by_term_ln = map(model, broom::tidy),
    perf_ln = map(model, broom::glance)
  )


by_term_ln <- model_ln %>%
  unnest(by_term_ln) %>%
  select(Keyword, term, estimate)
perf_ln <- model_ln %>%
  unnest(perf_ln) %>%
  select(Keyword, r.squared)

by_term_perf_ln <- by_term_ln %>%
  left_join(perf_ln, by = "Keyword") %>%
  arrange(desc(r.squared)) %>%
  dplyr::filter(term == "month_counter") %>%
  select(Keyword, rsq_log = r.squared, pct_log = estimate)
```

## Peaked

when the trend had a peaked and then went downwards

```{r}

peaked <- rolling %>%
  dplyr::filter(roll_avg_round > 0) %>%
  group_by(Keyword) %>%
  mutate(
    max = max(roll_avg_round),
    rmavg = mean(roll_avg_round),
    max_year = max(year),
    min_year = min(year)
  ) %>%
  slice_max(roll_avg_round, n = 1) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  mutate(ratio = rmavg / max) %>%
  # dplyr::filter(year != min_year, year != max_year) %>%
  select(Keyword, ratio)
```

## JOIN all



```{r}
# by_term_perf_ln
all_data <- kw_rsq_vol %>%
  left_join(by_term_perf_ln) %>%
  left_join(peaked) %>%
  # dplyr::filter(is.finite(r.squared), !is.na(estimate)) %>%
  dplyr::filter(if_all(c(avg_monthly_searches, estimate, r.squared, rsq_log, pct_log, ratio), ~ is.finite(.x))) %>%
  mutate(
    estimate = round(estimate),
    r.squared = round(r.squared, digits = 1),
    rsq_log = round(rsq_log, digits = 1),
    pct_log = round(pct_log, digits = 1),
    ratio = round(ratio, digits = 1)
  )
```

## save

```{r}
# write_rds(rolling, "rolling.rds")
# write_csv(all_data, "kw_rsq_vol.csv")
```

## visualise 

keywords with best R squared and positive coeffiecient values 
ie search queries with consistent and positive pattern of growth

```{r}

selection <- all_data %>%
  ungroup() %>%
  dplyr::filter(avg_monthly_searches > 100, estimate > 0) %>%
  slice_max(order_by = r.squared, n = 9, with_ties = FALSE) %>%
  pull(Keyword)

rolling %>%
  dplyr::filter(Keyword %in% selection) %>%
  ggplot(aes(x = year, y = roll_avg_round, colour = Keyword)) +
  geom_line() +
  facet_wrap(~Keyword, nrow = 3, scales = "free_y") +
  geom_smooth(method = "lm", se = F, linetype = 3, alpha = 0.3) +
  expand_limits(y = NULL) +
  theme(legend.position = "none", strip.text = element_text(size = 10)) +
  ylab("monthly_search_volume_USA")
```


## style plot for publication


```{r}

rolling %>%
  dplyr::filter(Keyword %in% selection) %>%
  plot_by_category(my_title = "", my_subtitle = "", time_col = "year")
```
